---
title: "Homework 1 - Building Blocks"
author: "Kristi Chau"
date: 2019-09-20
output: github_document
---

# Problem 1

**Creating a data frame**

```{r df_1, message = FALSE}
library(tidyverse)

HW1_df = tibble(
  HW1_samp = rnorm(8),
  HW1_samp_pos = HW1_samp > 0,
  HW1_char = c("This","is","my","first","home","work","assignment","yay"),
  HW1_factor = factor(c("Kittens","are","cute","Kittens","are","cute","are","cute")),
)
```

**Then trying to take the mean of each variable.**

```{r df_1_means, collapse = TRUE}
mean(pull(HW1_df, HW1_samp))
mean(pull(HW1_df, HW1_samp_pos))
mean(pull(HW1_df, HW1_char))
mean(pull(HW1_df, HW1_factor))
```

**What worked and what didn’t?**

_I was not able to take the mean of the last 2 variables because they were not numeric or logical vectors._ 

**Applying the as.numeric function to the logical, character, and factor variables**

```{r df_1_numeric, eval = FALSE}
as.numeric(pull(HW1_df, HW1_samp_pos))
as.numeric(pull(HW1_df, HW1_char))
as.numeric(pull(HW1_df, HW1_factor))
```

**What happens, and why?**

_The logical vector returned as numbers corresponding to TRUE or FALSE. The character vector was unable to return numeric values and instead returned NA. The factor vector returned numeric values although they corresponded to the levels rather than the vector itself._ 

# Problem 1.1

**Creating a new code chunk to manipulate variables from the dataframe:**

```{r df_1_convert, collapse = TRUE}
HW1_samp_num = as.numeric(pull(HW1_df, HW1_samp_pos))
HW1_samp_num * pull(HW1_df,HW1_samp)

HW1_samp_factor = as.factor(pull(HW1_df,HW1_samp_pos))
HW1_samp_factor * pull(HW1_df,HW1_samp)

HW1_samp_num2 = as.numeric(HW1_samp_factor)
HW1_samp_num2 * pull(HW1_df,HW1_samp)
```

# Problem 2

